---
title: "Project1PreProcessing"
author: "Elliot Schendel"
date: "10/8/2020"
output: word_document
---
## R Markdown Preprocessing Step 1: Remove Outliers

```{r}
#set working directory
setwd("C:/Users/Nathan/Desktop/DS110")

#read Group1EyeData dataset and attach it 
ProjectData=read.table("Group1EyeData.txt", sep = "\t", header = T)
attach(ProjectData)

#Dim function to count rows and columns
dim(ProjectData)

#look at summary...
summary(ProjectData)

#Making eyeDetect into factor variable
ProjectData$eyeDetect=as.factor(ProjectData$eyeDetect)

#finding column names
colnames(ProjectData)

#boxplot to visually see all outliers
boxplot(ProjectData)$out

######################################
#Round 1 of Outlier Removal
######################################

#Identifying AF3 outlier values. Remove the records which have those outliers from the entire dataset to make a new dataset without those outliers. Create a new boxplot to visualize the new dataset and compare dimensions of the old dataset to the new dataset.
af3val=boxplot(ProjectData$AF3)$out
af3val

rmaf3=which(AF3%in%af3val)
rmaf3

ProjectData1=ProjectData[-rmaf3,]

boxplot(ProjectData1)

dim(ProjectData)
dim(ProjectData1)

#Identifying F7 outlier values. Remove the records which have those outliers from the entire dataset to make a new dataset without those outliers. Create a new boxplot to visualize the new dataset and compare dimensions of the old dataset to the new dataset.
f7val=boxplot(ProjectData1$F7)$out
f7val

rmf7=which(ProjectData1$F7%in%f7val)
rmf7

ProjectData2=ProjectData1[-rmf7,]

boxplot(ProjectData2)

dim(ProjectData1)
dim(ProjectData2)

#Identifying F3 outlier values. Remove the records which have those outliers from the entire dataset to make a new dataset without those outliers. Create a new boxplot to visualize the new dataset and compare dimensions of the old dataset to the new dataset.
f3val=boxplot(ProjectData2$F3)$out
f3val

rmf3=which(ProjectData2$F3%in%f3val)
rmf3

ProjectData3=ProjectData2[-rmf3,]

boxplot(ProjectData3)

dim(ProjectData2)
dim(ProjectData3)

#Identifying FC5 outlier values. Remove the records which have those outliers from the entire dataset to make a new dataset without those outliers. Create a new boxplot to visualize the new dataset and compare dimensions of the old dataset to the new dataset.
fc5val=boxplot(ProjectData3$FC5)$out
fc5val

rmfc5=which(ProjectData3$FC5%in%fc5val)
rmfc5

ProjectData4=ProjectData3[-rmfc5,]

boxplot(ProjectData4)

dim(ProjectData3)
dim(ProjectData4)

#Identifying T. outlier values. Remove the records which have those outliers from the entire dataset to make a new dataset without those outliers. Create a new boxplot to visualize the new dataset and compare dimensions of the old dataset to the new dataset.
tval=boxplot(ProjectData4$T.)$out
tval

rmt=which(ProjectData4$T.%in%tval)
rmt

ProjectData5=ProjectData4[-rmt,]

boxplot(ProjectData5)

dim(ProjectData4)
dim(ProjectData5)

#Identifying P7 outlier values. Remove the records which have those outliers from the entire dataset to make a new dataset without those outliers. Create a new boxplot to visualize the new dataset and compare dimensions of the old dataset to the new dataset.
p7val=boxplot(ProjectData5$P7)$out
p7val

rmp7=which(ProjectData5$P7%in%p7val)
rmp7

ProjectData6=ProjectData5[-rmp7,]

boxplot(ProjectData6)

dim(ProjectData5)
dim(ProjectData6)

#Identifying O1 outlier values. Remove the records which have those outliers from the entire dataset to make a new dataset without those outliers. Create a new boxplot to visualize the new dataset and compare dimensions of the old dataset to the new dataset.
O1val=boxplot(ProjectData6$O1)$out
O1val

rmO1=which(ProjectData6$O1%in%O1val)
rmO1

ProjectData7=ProjectData6[-rmO1,]

boxplot(ProjectData7)

dim(ProjectData6)
dim(ProjectData7)

#Identifying O2 outlier values. Remove the records which have those outliers from the entire dataset to make a new dataset without those outliers. Create a new boxplot to visualize the new dataset and compare dimensions of the old dataset to the new dataset.
O2val=boxplot(ProjectData7$O2)$out
O2val

rmO2=which(ProjectData7$O2%in%O2val)
rmO2

ProjectData8=ProjectData7[-rmO2,]

boxplot(ProjectData8)

dim(ProjectData7)
dim(ProjectData8)

#Identifying P8 outlier values. Remove the records which have those outliers from the entire dataset to make a new dataset without those outliers. Create a new boxplot to visualize the new dataset and compare dimensions of the old dataset to the new dataset.
p8val=boxplot(ProjectData8$P8)$out
p8val

rmp8=which(ProjectData8$P8%in%p8val)
rmp8

ProjectData9=ProjectData8[-rmp8,]

boxplot(ProjectData9)

dim(ProjectData8)
dim(ProjectData9)

#Identifying T8 outlier values. Remove the records which have those outliers from the entire dataset to make a new dataset without those outliers. Create a new boxplot to visualize the new dataset and compare dimensions of the old dataset to the new dataset.
t8val=boxplot(ProjectData9$T8)$out
t8val

rmt8=which(ProjectData9$T8%in%t8val)
rmt8

ProjectData10=ProjectData9[-rmt8,]

boxplot(ProjectData10)

dim(ProjectData9)
dim(ProjectData10)

#Identifying FC6 outlier values. Remove the records which have those outliers from the entire dataset to make a new dataset without those outliers. Create a new boxplot to visualize the new dataset and compare dimensions of the old dataset to the new dataset.
fc6val=boxplot(ProjectData10$FC6)$out
fc6val

rmfc6=which(ProjectData10$FC6%in%fc6val)
rmfc6

ProjectData11=ProjectData10[-rmfc6,]

boxplot(ProjectData11)

dim(ProjectData10)
dim(ProjectData11)

#Identifying F4 outlier values. Remove the records which have those outliers from the entire dataset to make a new dataset without those outliers. Create a new boxplot to visualize the new dataset and compare dimensions of the old dataset to the new dataset.
f4val=boxplot(ProjectData11$F4)$out
f4val

rmf4=which(ProjectData11$F4%in%f4val)
rmf4

ProjectData12=ProjectData11[-rmf4,]

boxplot(ProjectData12)

dim(ProjectData11)
dim(ProjectData12)

#Identifying F8 outlier values. Remove the records which have those outliers from the entire dataset to make a new dataset without those outliers. Create a new boxplot to visualize the new dataset and compare dimensions of the old dataset to the new dataset.
f8val=boxplot(ProjectData12$F8)$out
f8val

rmf8=which(ProjectData12$F8%in%f8val)
rmf8

ProjectData13=ProjectData12[-rmf8,]

boxplot(ProjectData13)

dim(ProjectData12)
dim(ProjectData13)

#Identifying AF4 outlier values. Remove the records which have those outliers from the entire dataset to make a new dataset without those outliers. Create a new boxplot to visualize the new dataset and compare dimensions of the old dataset to the new dataset.
af4val=boxplot(ProjectData13$AF4)$out
af4val

rmaf4=which(ProjectData13$AF4%in%af4val)
rmaf4

ProjectData14=ProjectData13[-rmaf4,]

boxplot(ProjectData14)

dim(ProjectData13)
dim(ProjectData14)


######################################
#Round 2 of Outlier Removal
######################################


#Reidentifying AF3 outlier values. Remove the records which have those outliers from the entire dataset to make a new dataset without those outliers. Create a new boxplot to visualize the new dataset and compare dimensions of the old dataset to the new dataset.
AF3val=boxplot(ProjectData14$AF3)$out
AF3val

rmAF3=which(ProjectData14$AF3%in%AF3val)
rmAF3

ProjectData15=ProjectData14[-rmAF3,]

boxplot(ProjectData15)

dim(ProjectData14)
dim(ProjectData15)

#Reidentifying F7 outlier values. Remove the records which have those outliers from the entire dataset to make a new dataset without those outliers. Create a new boxplot to visualize the new dataset and compare dimensions of the old dataset to the new dataset.
F7val=boxplot(ProjectData15$F7)$out
F7val

rmF7=which(ProjectData15$F7%in%F7val)
rmF7

ProjectData16=ProjectData15[-rmF7,]

boxplot(ProjectData16)

dim(ProjectData15)
dim(ProjectData16)

#Reidentifying F3 outlier values. Remove the records which have those outliers from the entire dataset to make a new dataset without those outliers. Create a new boxplot to visualize the new dataset and compare dimensions of the old dataset to the new dataset.
F3val=boxplot(ProjectData16$F3)$out
F3val

rmF3=which(ProjectData16$F3%in%F3val)
rmF3

ProjectData17=ProjectData16[-rmF3,]

boxplot(ProjectData17)

dim(ProjectData16)
dim(ProjectData17)

#Reidentifying FC5 outlier values. Remove the records which have those outliers from the entire dataset to make a new dataset without those outliers. Create a new boxplot to visualize the new dataset and compare dimensions of the old dataset to the new dataset.
FC5val=boxplot(ProjectData16$FC5)$out
FC5val

rmFC5=which(ProjectData17$FC5%in%FC5val)
rmFC5

ProjectData18=ProjectData17[-rmFC5,]

boxplot(ProjectData18)

dim(ProjectData17)
dim(ProjectData18)

#Reidentifying T. outlier values. Remove the records which have those outliers from the entire dataset to make a new dataset without those outliers. Create a new boxplot to visualize the new dataset and compare dimensions of the old dataset to the new dataset.
T.val=boxplot(ProjectData18$T.)$out
T.val

rmT.=which(ProjectData18$T.%in%T.val)
rmT.

ProjectData19=ProjectData18[-rmT.,]

boxplot(ProjectData19)

dim(ProjectData18)
dim(ProjectData19)

#Reidentifying P7 outlier values. Remove the records which have those outliers from the entire dataset to make a new dataset without those outliers. Create a new boxplot to visualize the new dataset and compare dimensions of the old dataset to the new dataset.
P7val=boxplot(ProjectData19$P7)$out
P7val

rmP7=which(ProjectData19$P7%in%P7val)
rmP7

ProjectData20=ProjectData19[-rmP7,]

boxplot(ProjectData20)

dim(ProjectData19)
dim(ProjectData20)

#Reidentifying 01 outlier values. Remove the records which have those outliers from the entire dataset to make a new dataset without those outliers. Create a new boxplot to visualize the new dataset and compare dimensions of the old dataset to the new dataset.
O1val=boxplot(ProjectData20$O1)$out
O1val

rmO1=which(ProjectData20$O1%in%O1val)
rmO1

ProjectData21=ProjectData20[-rmO1,]

boxplot(ProjectData21)

dim(ProjectData20)
dim(ProjectData21)

#Reidentifying 02 outlier values. Remove the records which have those outliers from the entire dataset to make a new dataset without those outliers. Create a new boxplot to visualize the new dataset and compare dimensions of the old dataset to the new dataset.
O2val=boxplot(ProjectData21$O2)$out
O2val

rmO2=which(ProjectData21$O2%in%O2val)
rmO2

ProjectData22=ProjectData21[-rmO2,]

boxplot(ProjectData22)

dim(ProjectData21)
dim(ProjectData22)

#Reidentifying P8 outlier values. Remove the records which have those outliers from the entire dataset to make a new dataset without those outliers. Create a new boxplot to visualize the new dataset and compare dimensions of the old dataset to the new dataset.
P8val=boxplot(ProjectData22$P8)$out
P8val

rmP8=which(ProjectData22$P8%in%P8val)
rmP8

ProjectData23=ProjectData22[-rmP8,]

boxplot(ProjectData23)

dim(ProjectData22)
dim(ProjectData23)

#Reidentifying T8 outlier values. Remove the records which have those outliers from the entire dataset to make a new dataset without those outliers. Create a new boxplot to visualize the new dataset and compare dimensions of the old dataset to the new dataset.
T8val=boxplot(ProjectData23$T8)$out
T8val

rmT8=which(ProjectData23$T8%in%T8val)
rmT8

ProjectData24=ProjectData23[-rmT8,]

boxplot(ProjectData24)

dim(ProjectData23)
dim(ProjectData24)

#Reidentifying FC6 outlier values. Remove the records which have those outliers from the entire dataset to make a new dataset without those outliers. Create a new boxplot to visualize the new dataset and compare dimensions of the old dataset to the new dataset.
FC6val=boxplot(ProjectData24$FC6)$out
FC6val

rmFC6=which(ProjectData24$FC6%in%FC6val)
rmFC6

ProjectData25=ProjectData24[-rmFC6,]

boxplot(ProjectData25)

dim(ProjectData24)
dim(ProjectData25)

#Reidentifying F4 outlier values. Remove the records which have those outliers from the entire dataset to make a new dataset without those outliers. Create a new boxplot to visualize the new dataset and compare dimensions of the old dataset to the new dataset.
F4val=boxplot(ProjectData25$F4)$out
F4val

rmF4=which(ProjectData25$F4%in%F4val)
rmF4

ProjectData26=ProjectData25[-rmF4,]

boxplot(ProjectData26)

dim(ProjectData25)
dim(ProjectData26)

#Reidentifying F8 outlier values. Remove the records which have those outliers from the entire dataset to make a new dataset without those outliers. Create a new boxplot to visualize the new dataset and compare dimensions of the old dataset to the new dataset.
F8val=boxplot(ProjectData26$F8)$out
F8val

rmF8=which(ProjectData26$F8%in%F8val)
rmF8

ProjectData27=ProjectData26[-rmF8,]

boxplot(ProjectData27)

dim(ProjectData26)
dim(ProjectData27)

#Reidentifying AF4 outlier values. Remove the records which have those outliers from the entire dataset to make a new dataset without those outliers. Create a new boxplot to visualize the new dataset and compare dimensions of the old dataset to the new dataset.
AF4val=boxplot(ProjectData27$AF4)$out
AF4val

rmAF4=which(ProjectData27$AF4%in%AF4val)
rmAF4

ProjectData28=ProjectData27[-rmAF4,]

boxplot(ProjectData28)

dim(ProjectData27)
dim(ProjectData28)

######################################
#Round 3 of Outlier Removal
######################################

#Reidentifying AF3 outlier values. Remove the records which have those outliers from the entire dataset to make a new dataset without those outliers. Create a new boxplot to visualize the new dataset and compare dimensions of the old dataset to the new dataset.
af3val2=boxplot(ProjectData28$AF3)$out
af3val2

rmaf3v2=which(ProjectData28$AF3%in%af3val2)
rmaf3v2

ProjectData29=ProjectData28[-rmaf3v2,]

boxplot(ProjectData29)

dim(ProjectData28)
dim(ProjectData29)

#Reidentifying F7 outlier values. Remove the records which have those outliers from the entire dataset to make a new dataset without those outliers. Create a new boxplot to visualize the new dataset and compare dimensions of the old dataset to the new dataset.
f7val2=boxplot(ProjectData29$F7)$out
f7val2

rmf7v2=which(ProjectData29$F7%in%f7val2)
rmf7v2

ProjectData30=ProjectData29[-rmf7v2,]

boxplot(ProjectData30)

dim(ProjectData29)
dim(ProjectData30)

#Reidentifying F3 outlier values. Remove the records which have those outliers from the entire dataset to make a new dataset without those outliers. Create a new boxplot to visualize the new dataset and compare dimensions of the old dataset to the new dataset.
f3val2=boxplot(ProjectData30$F3)$out
f3val2

#We have no outliers for 'F3', moving to next variable with visible outliers
boxplot(ProjectData30)
#boxplot(ProjectData30$AF3)
#boxplot(ProjectData30$F7)
#boxplot(ProjectData30$T.)
#boxplot(ProjectData30$P7)
#boxplot(ProjectData30$T8)
#boxplot(ProjectData30$FC6)

#Reidentifying T. outlier values. Remove the records which have those outliers from the entire dataset to make a new dataset without those outliers. Create a new boxplot to visualize the new dataset and compare dimensions of the old dataset to the new dataset.
tval2=boxplot(ProjectData30$T.)$out
tval2

rmtv2=which(ProjectData30$T.%in%tval2)
rmtv2

ProjectData31=ProjectData30[-rmtv2,]

boxplot(ProjectData31)

#Reidentifying P7 outlier values. Remove the records which have those outliers from the entire dataset to make a new dataset without those outliers. Create a new boxplot to visualize the new dataset and compare dimensions of the old dataset to the new dataset.
p7val2=boxplot(ProjectData31$P7)$out
p7val2

rmp7v2=which(ProjectData31$P7%in%p7val2)
rmp7v2

ProjectData32=ProjectData31[-rmp7v2,]

boxplot(ProjectData32)

dim(ProjectData31)
dim(ProjectData32)

#Reidentifying T8 outlier values. Remove the records which have those outliers from the entire dataset to make a new dataset without those outliers. Create a new boxplot to visualize the new dataset and compare dimensions of the old dataset to the new dataset.
t8val2=boxplot(ProjectData32$T8)$out
t8val2

rmt8v2=which(ProjectData32$T8%in%t8val2)
rmt8v2

ProjectData33=ProjectData32[-rmt8v2,]

boxplot(ProjectData33)

dim(ProjectData32)
dim(ProjectData33)

#Reidentifying FC6 outlier values. Remove the records which have those outliers from the entire dataset to make a new dataset without those outliers. Create a new boxplot to visualize the new dataset and compare dimensions of the old dataset to the new dataset.
fc6val2=boxplot(ProjectData33$FC6)$out
fc6val2

rmfc6v2=which(ProjectData33$FC6%in%fc6val2)
rmfc6v2

ProjectData34=ProjectData33[-rmfc6v2,]

boxplot(ProjectData34)

dim(ProjectData33)
dim(ProjectData34)

#Reidentifying AF4 outlier values. Remove the records which have those outliers from the entire dataset to make a new dataset without those outliers. Create a new boxplot to visualize the new dataset and compare dimensions of the old dataset to the new dataset.
af4val2=boxplot(ProjectData34$AF4)$out
af4val2

rmaf4v2=which(ProjectData34$AF4%in%af4val2)
rmaf4v2

ProjectData35=ProjectData34[-rmaf4v2,]

boxplot(ProjectData35)

dim(ProjectData34)
dim(ProjectData35)

#boxplot(ProjectData35$AF3)
#boxplot(ProjectData35$F7)
#boxplot(ProjectData35$F8)

######################################
#Round 4 of Outlier Removal
######################################

#Reidentifying AF3outlier values. Remove the records which have those outliers from the entire dataset to make a new dataset without those outliers. Create a new boxplot to visualize the new dataset and compare dimensions of the old dataset to the new dataset.
af3val3=boxplot(ProjectData35$AF3)$out
af3val3

rmaf3v3=which(ProjectData35$AF3%in%af3val3)
rmaf3v3

ProjectData36=ProjectData35[-rmaf3v3,]

boxplot(ProjectData36)

#Reidentifying F7 outlier values. Remove the records which have those outliers from the entire dataset to make a new dataset without those outliers. Create a new boxplot to visualize the new dataset and compare dimensions of the old dataset to the new dataset.
f7val3=boxplot(ProjectData36$F7)$out
f7val3

rmf7v3=which(ProjectData36$F7%in%f7val3)
rmf7v3

ProjectData37=ProjectData36[-rmf7v3,]

boxplot(ProjectData37)
#boxplot(ProjectData37$O1)
#boxplot(ProjectData37$F8)

#Reidentifying o1 outlier values. Remove the records which have those outliers from the entire dataset to make a new dataset without those outliers. Create a new boxplot to visualize the new dataset and compare dimensions of the old dataset to the new dataset.
O1val2=boxplot(ProjectData37$O1)$out
O1val2

rmO1v2=which(ProjectData37$O1%in%O1val2)
rmO1v2

ProjectData38=ProjectData37[-rmO1v2,]

boxplot(ProjectData38)

#Reidentifying F8 outlier values. Remove the records which have those outliers from the entire dataset to make a new dataset without those outliers. Create a new boxplot to visualize the new dataset and compare dimensions of the old dataset to the new dataset.
f8val2=boxplot(ProjectData38$F8)$out
f8val2

rmf8v3=which(ProjectData38$F8%in%f8val2)
rmf8v3

ProjectData39=ProjectData38[-rmf8v3,]

boxplot(ProjectData39)
dim(ProjectData39)

######################################
#Final check for outlier values
######################################

#Checking all variables for any remaining outliers
#boxplot(ProjectData39$AF3)
#boxplot(ProjectData39$F7)
#boxplot(ProjectData39$F3)
#boxplot(ProjectData39$FC5)
#boxplot(ProjectData39$T.)
#boxplot(ProjectData39$P7)
#boxplot(ProjectData39$O1)
#boxplot(ProjectData39$O2)
#boxplot(ProjectData39$P8)
#boxplot(ProjectData39$T8)
#boxplot(ProjectData39$FC6)
#boxplot(ProjectData39$F4)
#boxplot(ProjectData39$F8)
#boxplot(ProjectData39$AF4)

#Outlier check complete. 'ProjectData39' the new data table with 3257 records and 15 rows of outlier free data.

#Write the outlier free dataset into a txt file into working directory
write.table(ProjectData39, "Group1EyeDetectDataNoOutliers.txt", sep = "\t", row.names = F, col.names = TRUE)
```

## R Markdown Preprocessing Step 2: Creating 'Train' and 'Test' datasets from 'ProjectData39'

```{r}
#read Group1EyeDataNoOutliers dataset and attach it 
OutlierFreeProjectData=read.table("Group1EyeDetectDataNoOutliers.txt", sep = "\t", header = T)
attach(OutlierFreeProjectData)

#Reading dimensions/summary of OutlierFreeProjectData.Using boxplot to prove this data is outlier free.
summary(OutlierFreeProjectData)
boxplot(OutlierFreeProjectData)
dim(OutlierFreeProjectData)
#As stated above, this new dataset of outlier free value has 3257 rows of records.

#Creating a function to remove 80% of the values from OutlierFreeProjectData without replacement.80% of 3257 is 2,605.6 (For simplicity, we will round this up to 2,606 and extract this amount).
RandomRows=sample(1:3257, 2606, replace = FALSE)

#Creating 'Train' dataset by extracting these rows from the OutlierFreeProjectData
Group1EyeDetectTRAIN=OutlierFreeProjectData[RandomRows, ]

#Creating 'Test' dataset by removing these records from the OutlierFreeProjectData
Group1EyeDetectTEST=OutlierFreeProjectData[-RandomRows, ]

#Confirming sizes of Train and Test datasets
dim(Group1EyeDetectTRAIN)
dim(Group1EyeDetectTEST)

#Writing Train and Test datasets into txt files into working directory
write.table(Group1EyeDetectTRAIN, "EyeDetect-TRAIN.txt", sep = "\t", row.names = F, col.names = TRUE)
write.table(Group1EyeDetectTEST, "EyeDetect-TEST.txt", sep = "\t", row.names = F, col.names = TRUE)
```

## R Markdown Preprocessing Step 3: Applying PCA to 'EyeDetect-TRAIN.txt' dataset

```{r}
#read Group1EyeDetect-TRAIN dataset and attach it 
EyeDetectTRAIN=read.table("EyeDetect-TRAIN.txt", sep = "\t", header = T)
attach(EyeDetectTRAIN)

#head/summary/column names of TRAIN
head(EyeDetectTRAIN)
summary(EyeDetectTRAIN)
colnames(EyeDetectTRAIN)

#Apply PCA to reduce the number of variables that can provide the maximum variabilty in the dataset
eyePCA=prcomp(EyeDetectTRAIN[,-15], scale. = TRUE, center = TRUE)
eyePCA

summary(eyePCA)


plot(eyePCA,type="l",pch=16)
#based on cumulative proportion we select 8 principal components
# we then used r>|.49| and we will select the variables F7, FC5, T, O1, F8
# using r>|.45| we will select variable T., P7, O1, FC5, F7, FC6, F8

colnames(EyeDetectTRAIN)

#Now we will remove all predictor variables besides the ones we selected in our PCA analysis
PCAEyeDetectTRAIN <- EyeDetectTRAIN[,-c(1,3,8,9,10,12,14)]
```

## R Markdown Preprocessing Step 4: Data Descriptions: Numerical Description of Variables

```{r}

head(EyeDetectTRAIN)
str(EyeDetectTRAIN)

#Let's Provide Numerical Summaries of each of the Variables that remain after PCA. Recall the variables that remain after PCA are T., P7, O1, FC5, F7, FC6, F8

#T.
summary(T.)

#P7
summary(P7)

#O1
summary(O1)

#FC5
summary(FC5)

#F7
summary(F7)

#FC6
summary(FC6)

#F8
summary(F8)

#Eye Detect
table(EyeDetectTRAIN$eyeDetect)




```


## R Markdown Preprocessing Step 5: Data Descriptions: Visualization of Variables

```{r}
#Let's Start visualizing each of the Variables that remain after PCA. Recall the variables that remain after PCA are T., P7, O1, FC5, F7, FC6, F8

#In order to visualize the data, we need to first require ggplot2
require(ggplot2)

#Before we start, lets convert eyeDetect to a factor variable to make it easier to use within the data visualization.
summary(eyeDetect)
EyeDetectTRAIN$eyeDetect=as.factor(EyeDetectTRAIN$eyeDetect)
summary(eyeDetect)

#We will begin the visualization by providing a simple histogram of each variable to get an idea of how they are distributed.


#ggplot(Housing, aes(x=Land.Value, y=Home.Value))
#ggplot(Housing, aes(x=Land.Value, y=Home.Value))+geom_point()

#T.
ggplot(EyeDetectTRAIN, aes(x=T.))+geom_histogram(col="white", fill="red", bins = 25) + ggtitle("Distrubition of T.")

#P7
ggplot(EyeDetectTRAIN, aes(x=P7))+geom_histogram(col="white", fill="orange", bins = 25) + ggtitle("Distrubition of P7")

#O1
ggplot(EyeDetectTRAIN, aes(x=O1))+geom_histogram(col="white", fill="yellow", bins = 25) + ggtitle("Distrubition of O1")

#FC5
ggplot(EyeDetectTRAIN, aes(x=FC5))+geom_histogram(col="white", fill="green", bins = 25) + ggtitle("Distrubition of FC5")

#F7
ggplot(EyeDetectTRAIN, aes(x=F7))+geom_histogram(col="white", fill="cyan", bins = 25) + ggtitle("Distrubition of F7")

#FC6
ggplot(EyeDetectTRAIN, aes(x=FC6))+geom_histogram(col="white", fill="blue", bins = 25) + ggtitle("Distrubition of FC6")

#F8
ggplot(EyeDetectTRAIN, aes(x=F8))+geom_histogram(col="white", fill="purple", bins = 25) + ggtitle("Distrubition of F8")

#Now let's do a set of multivariate plots. Let's see how the eyeDetect variable changes with each variable through layering of histograms. This will give us an idea of how EyeDetect changes as each of the variables increase or decrease.

#T.
ggplot(data=EyeDetectTRAIN, aes(x=T., fill=eyeDetect))+geom_histogram() + ggtitle("Distrubition of eyeDetect over T.")

#P7
ggplot(data=EyeDetectTRAIN, aes(x=P7, fill=eyeDetect))+geom_histogram() + ggtitle("Distrubition of eyeDetect over P7")

#O1
ggplot(data=EyeDetectTRAIN, aes(x=O1, fill=eyeDetect))+geom_histogram() + ggtitle("Distrubition of eyeDetect over O1")

#FC5
ggplot(data=EyeDetectTRAIN, aes(x=FC5, fill=eyeDetect))+geom_histogram() + ggtitle("Distrubition of eyeDetect over FC5")

#F7
ggplot(data=EyeDetectTRAIN, aes(x=F7, fill=eyeDetect))+geom_histogram() + ggtitle("Distrubition of eyeDetect over F7")

#FC6
ggplot(data=EyeDetectTRAIN, aes(x=FC6, fill=eyeDetect))+geom_histogram() + ggtitle("Distrubition of eyeDetect over FC6")

#F8
ggplot(data=EyeDetectTRAIN, aes(x=F8, fill=eyeDetect))+geom_histogram() + ggtitle("Distrubition of eyeDetect over F8")

#From these layered histograms, it seems that for some variables, eyeDetect has a more clear relationship with the given variable (P7, O1), while for the others it seems to follow a relatively normally distribution. Let's explore the variables which seem to have a more clear relationship with eyeDetect further. These variables are P7 & O1. Let's get an idea of how the eyedetect variable changes when P7 & O1 are compared to a variable for which eyeDetect is roughly normally distributed (T.).


#P7 vs T.
ggplot(data=EyeDetectTRAIN, aes(y=P7,x=T., col=eyeDetect))+geom_point() + ggtitle("Rate of eyeDetect on P7 vs T.") +
  xlab("T.") + ylab("P7")


#From this chart we can see that as P7 increases (y axis), the rate of eyeDetect decreases (goes from blue to red). Most of the observations where eyeDetect is 1 are focused in the lower to middle values of P7.

#O1 vs T.
ggplot(data=EyeDetectTRAIN, aes(y=O1,x=T., col=eyeDetect))+geom_point() + ggtitle("Rate of eyeDetect on O1 vs T.") +
  xlab("T.") + ylab("O1")

#From this chart we can see the relationship between O1 (y axis) and eyeDetect. As O1 nears the lower and upper bounds of its values, eyeDetect is mostly 1 (Blue). As it nears the middle of the distrubtion, an eyeDetect value of 0 (red) becomes the trend.






```

